An Abstraction-based Approach to Network Synthesis

### 2016-11-02 paper/abstraction/paper.tex
intro edits

\newcommand{\sysname}{{\small \sf Methane}\xspace}
\newcommand{\sysnamesec}{{\sf Methane}\xspace}

@@

@@

>>>>

\newcommand{\sysname}{{\small \sf Methane}\xspace}
\newcommand{\sysnamesec}{{\sf Methane}\xspace}
\newcommand{\propane}{{\small \sf Propane}\xspace}

@@

\section{Outline/Notes}

@@

% Hence, one approach to improving network
% reliability is to develop better languages for network configuration.
% Such languages should be defined at a higher level of abstraction than
% existing configuration languages and provide support built-in support for
% verification of important properties, thereby avoiding entire classes
% of configuration errors automatically.

An important class of networks are the large \emph{structured networks}
commonly deployed in industrial data centers.  These networks run
many critical services and every second of downtime is at best
costly and at worst dangerous.  Unfortunately, it is an enormous 
challenge to keep these complex networks up and
running 24/7~\cite{mahajan+:bgp-misconfiguration,feamster+:rcc,batfish,dc-failure-study}.
Moreover, while hardware faults, rodents, power failures and natural
disasters can all cause significant downtime, studies have shown that
\emph{human error}, which occurs during network configuration or
update, is the single large cause of these
outages~\cite{juniper-study}.

% These networks often
% host a myriad of services and every second of downtime imposes significant
% financial cost or disruption to consumers.  
Such networks typically have uniform, tree-shaped topologies that involve
hundreds or even thousands of devices.  Since it is not possible for humans
to think about configuring each individual device independently, 
network operators typically classify the devices, assigning them a particular
\emph{role} such as ``top-of-rack switch'' or ``tier-one aggregation switch.''
Each such role is codified as a \emph{configuration template}---essentially
a macro that may be instantiated with different concrete values to
generate a configuration for each device that plays the role.

Such templates have no particular semantics other than the semantics
that one obtains from instantiating a template's parameters.  Moreover,
template construction is a manual process:  network engineers do their
best to design templates by hand that will work effectively in the context 
of their particular network topology.  However, there is no support for
validating key properties such as connectivity and fault tolerance, 
aside from instantiating the templates and testing.  In addition,
network engineers have no tools to help them design for network
evolution.  For instance, the process of adding new equipment to
expand the capacity of data center may not be compatible with the original
templates, or may cause a trickle-down of changes to be made network-wide.
Unfortunately, each change to a router configuration typically incurs
some risk of temporary instable in a network, so minimizing such changes
is an important design goal.  It is no wonder that a major
network outage, such as the one that grounded all United Air flights
in July 2015~\cite{x}, makes international news every few months.

To combat these problems, in this paper, we present \sysname, a 
new high-level language designed to 
provide improved support for the configuration of large, structured networks.
\sysname allows network operators to define abstract network topologies that
classify devices in terms of abstract roles and specify the connectivity
invariants satisfied by each role.  For instance, a programmer may
define separate roles for ``top-of-rack switch'' and 
``tier-one aggregation switch'' and specify that every top-of-rack switch
is connected to at least two different tier-one aggregation switches.

In order to specify \emph{routing policy} for such abstract networks,
\sysname borrows notation from the recently-defined \propane
programming language~\cite{propane}.  However, whereas \propane policies
refer to individual devices, \sysname policies allows users to
refer to abstract roles.  The \sysname compiler compiles these role-based
policies into templates, with one template per role, and these templates
may be instantiated to generate concrete configurations.

During compilation, the \sysname compiler also analyzes policy with respect to the
abstract topology and guarantees that the configurations will 
faithfully implement the high-level policy on \emph{any} concrete
topology that satisfies the given connectivity invariants.  The compiler
will also optionally perform a second fault tolerance analysis that 
determines the minimum number of link failures required to disconnect
\emph{any} concrete topology that once again satisfies the connectivity
invariants.

Finally, generated \sysname configurations are designed to
be \emph{evolution-friendly} in the sense that if an operator adds or
removes any single link or device, they are guaranteed that only 
the configurations of adjacent devices need change (for instance, to
activate a port on which the new device is connected).  Doing so
helps minimize network instability as data centers are expanded or devices are
temporarily taken offline for maintenance.

As an aside, a pleasant side effect of this new compiler
infrastructure, designed to make life easier on network operators, is
that for structured networks, it scales substantially better than its
competitor \propane does: as the number of individual devices in a
network grows, the number of distinct roles will often stay constant.
Consequently, the \sysname compiler may be two orders of magnitude
faster than \propane for large structured networks.

To summarize, the central contribution of this paper are:

\begin{enumerate}
\item Design of new network topology abstraction mechanisms involving roles and
connectivity invariants.
\item Design of two new algorithms for analyzing abstract networks: one which
guarantees correct compilation and one which determines fault tolerance 
properties of network policy.  Both algorithms provide guarantees for any 
concrete network satisfying the abstract connectivity invariants.
\item Design of template generation algorithms that support network
evolution with minimal change to existing devices.
\item Implementation and evaluation of a \sysname compiler that generates 
configuration templates and is up to two orders of magnitude faster 
than its closest competitor on large networks.
\end{enumerate}

%% Indeed, over the last several years, many researchers have begun to
%% develop such languages~\cite{frenetic,pyretic,pane,nettle,netkat}.
%% However, the bulk of this research has targeted software-defined
%% networks (SDN), a new networking technology that has emerged over the last
%% decade.  Unfortunately, the vast majority of today's networks
%% continue to use traditional networking techology---upgrading to SDN is
%% expensive and technically challenging.

%% In a traditional network, every router runs an independent instance of 
%% a distributed protocol that
%% exchange messages about available routes to destination addresses.  
%% Each router is configured independently, but the overall behavior
%% of the network---the paths from one server to another, for instance---depend
%% upon the
